The LSST-DESC 3x2pt Tomography Optimization Challenge

### 2021-09-27 paper/paper.tex
changes for referee report

Various recent work have explored general tomographic binning strategies. The general question of optimization
was recently discussed in detail
in \citet{rainbow} using a self-organizing map (SOM) approach to target up to five tomographic bins.  For this
configuration they find that equally spaced redshift bins are a good proxy
for an optimized method, but note, importantly, that we should not be constrained to trying to directly
match our bin selections to specific tomographic bin edges. 
They also highlight the utility of rejecting outlier or hard-to-classify objects from samples, and the ability of SOMs
to do so cleanly.
\citet{taylor18} considered using
fine tomographic bins as an alternative to fully 3D lensing, and found that the strategy of equal numbers
per bin is less effective at high redshift, a result we will echo in this paper. The fine-binning strategy
also allows a set of nulling methods to be applied, offering various advantages in avoiding poorly-understood
regimes \citep{taylor18b,bnt,xcut}.
Most recently, \citet{euclid-nz} studied this question with respect to the Euclid mission, using the Flagship
simulation \citep{euclid-flagship} and applying realistic photometric redshift estimation methods to the
sample (we defer this question in this work). They highlight the power of using large numbers of bins,
differences in ideal binnings when
changing the choice of data (the inclusion galaxy galaxy-lensing), and
note the value in discarding galaxies with high redshift uncertainty. We will echo many of these issues below.

@@

\begin{table*}[]
	\begin{tabular}{|l|p{8cm}|c|}
	\hline
	\textbf{Name} & \textbf{Description} & \textbf{Submitters} \\
	\hline
	\textsc{Random} & Random assignment (control method)  & Organizers \\
	\textsc{IBandOnly} & i-band only (control method)  & Organizers \\
	\textsc{RandomForest} & A collection of individual decision trees & Organizers \\
	\textsc{LSTM} & Long Short-Term Memory neural network & CRB, BF, GT, ESC, EJG \\
	\textsc{AutokerasLSTM}\tablenotemark{a} & Automaticaly configured LSTM variant. & CRB, BF, GT, ESC, EJG \\
	\textsc{LGBM} & Light Gradient Boosted Machine & CRB, BF, GT, ESC, EJG \\
	\textsc{TCN} & Temporal Convolution Network & CRB, BF, GT, ESC, EJG \\
	\textsc{CNN} & Convolutional Neural Network & CRB, BF, GT, ESC, EJG \\
	\textsc{Ensemble} & Combination of three other network methods & CRB, BF, GT, ESC, EJG \\
	\textsc{SimpleSOM} & Self-Organizing Map & AHW \\
	\textsc{PQNLD} & Extension of {\sc SimpleSom} combining with template-based redshift estimation & AHW \\
	\textsc{UTOPIA} & Nearest-neighbours, optimised in the limit of large representative training sets & AHW \\
	\textsc{ComplexSOM} & Extension of {\sc SimpleSom} with an additional assignment optimization & DA, AHW, AN, EG, BA, ABr \\
	\textsc{GPzBinning} & Gaussian Process redshift estimator and binner & PH \\
	\textsc{JaxCNN/JaxResnet} & Two related CNN-based bin edge optimizers & AP \\
	\textsc{NeuralNetwork1/2} & Dense network optimizing bin assignment for two metrics & FL \\
	\textsc{PCACluster} & Principal Component Analysis of fluxes then optimized clustering & DG \\
	\textsc{ZotBin/ZotNet} & Two related neural network methods with a preprocesing normalizing flow & DK \\
	\textsc{FFNN} & Feed-forward neural network & EN \\
	\textsc{FunBins} & Random forest with various nominal bin edge selection & ABa \\
	\textsc{MLPQNA} & Multi-layer perceptron & SC, MB \\
	\textsc{Stacked Generalization} & Combination of Gradient Boosted Trees classifiers with standard Random Forests (50/50) & JEC\\
	\hline
	\end{tabular}
	\caption{Methods entered into the challenge. The algorithms are described more fully in \autoref{app:methods}.}
	\label{tab:entrants}
	\tablenotetext{a}{A technical problem, involving conflicting versions of libraries and incompatible hardware, meant that AutoKerasLSTM could not be run in the final challenge.}
\end{table*}

@@

\begin{table*}[]
	\begin{tabular}{|l|llll|llll|}
		\hline
		& \multicolumn{4}{c|}{\textbf{riz}}      & \multicolumn{4}{c|}{\textbf{griz}}                                \\ \hline
		\textbf{Method} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} \\ \hline
		\input{results/tables/table_FOM_DETF_3x2_dc2.tex}
		\hline
	\end{tabular}
	\caption{Values of the 3x2pt DETF ($w_0,w_a$) figure-of-merit achieved by entrant methods on the 
		CosmoDC2 part of the challenge. The horizontal line separates two general approaches to the problem, as discussed in \autoref{sec:train-vs-fix}: those below the line used fixed
		target edges for tomographic bins and optimized the assignment of objects to those bins,
		whereas those above also optimized the target edges themselves.}
	\label{tab:cosmodc2}
\end{table*}

@@

\begin{figure}[htbp]
	% \centering
	\includegraphics[width=1\linewidth]{results/nzbinned_dc2_funbins.pdf}
	\caption{The tomographic $n(z)$ values generated by the winning method in the challenge, {\sc funbins},
		generating nine bins with the CosmoDC2 data.
		The upper panel shows results for the method applied to $riz$ bands, and the middle on $griz$.
		The lower panel is for comparison and is not part of the challenge;
		it illustrates {\sc funbins} results using all the $ugrizy$ bands.
		The increased spread and overlap of the bins as bands are removed decreases the overall constraining
		power and signal-to-noise.
	}
	\label{fig:funbin_nz}
\end{figure}

No one method dominates the metrics across the different scenarios considered.
Before the challenge began we somewhat arbitrarily chose 
the 9-bin 3x2pt $griz$ CosmoDC2 $w_0-w_a$ metric as the fiducial scenario on which
an overall winning method would be selected.
By this criterion the best method was, by a small margin, \textsc{funbins},
which used a random forest to assign galaxies to after choosing nominal bin edges that split 
the range spanned
by the sample into equal co-moving distance bins. 
As described in \autoref{sec:funbins},  this splitting was designed
to minimze the shot noise for the angular power in each bin, and this has proven successful. 
Figure \ref{fig:funbin_nz} shows the number density obtained by this method, for both the riz and griz 
scenarios, and an additional post-challenge run using the full Rubin $ugrizy$ bands.
The metrics \textsc{funbins}
obtained on $ugrizy$ are barely higher than those with $griz$ -- the largest
increase is 8\% for the $\Omega_c-\sigma_8$ metric, with the remainder all 5\% or less.

@@

\begin{figure*}[htbp]
% \centering
\includegraphics[width=0.9\linewidth]{results/nzoverlap_dc2_funbins.pdf}
\caption{The pairwise fractions $f_\mathrm{mis}$ of total objects in redshift regions where multiple tomographic bins contain galaxies, for 
the \textsc{funbins} tomography method and three choices of bands.  
The $x$- and $y$-axes 
show tomographic bin indices, 
and the colour represents the fraction of objects (compared to the total number in the challenge) that are 
at redshifts where both tomographic bins have members, in bins of width $\delta z=10^{-3}$.  
This metric corresponds to
the fraction of the area that is shaded in two colours in Figure \ref{fig:funbin_nz}.
The total-overlap figure gives the total fraction of objects
that are ambiguous over all bins, and is the sum of all the pairwise values.}
\label{fig:overlap}
\end{figure*}

\textsc{funbins} seems to be a good and stable general method, though several other methods perform as well
or slightly better in other scenarios.  Notably, many methods reach the same plateau in lensing-only metrics,
suggesting this is somewhat easier to optimize; this is consistent with the general reason that photometric
redshifts suffice for lensing: the broad lensing kernels are less sensitive to small changes.

@@

\subsection{Non-assignment}
The challenge did not require entrants to assign all galaxies to bins; aside from the impact of reduced final sample size on the metrics, no additional penalty was applied for discarding galaxies entirely.  Several methods
took advantage of this.  The ComplexSom method (see \autoref{sec:csom}) found
it did not improve scores, but as noted in \autoref{sec:pqnld} this may be an artifact
of the high density of the training data, and so could be explored in future challenges\footnote{In real data or more realistic simulations discarding galaxies has more complicated effects than simply decreasing sample size as it does here - removing objects with broad $p(z)$ can narrow the final $n(z)$, not just lower it. This was not a factor in this challenge because we use true redshifts to calculate our metrics.}. See \citet{euclid-flagship} for a discussion of this topic in the Euclid context.

@@

The different metrics are generally well-correlated (this also holds true for the metrics not 
shown here), especially at the highest values of the metrics representing the most successful
methods. This is particularly so when comparing the Buzzard and CosmoDC2 metrics, which is encouraging 
as it implies these methods are broadly stable to reasonable changes in the underlying galaxy
distribution, and hence on the survey properties.  The largest outlier comes from the 
{\sc Neural Network} entry.

@@

The value of the extra FoM that could be gained should be weighed against 
the challenge of high-accuracy
PSF measurement for this band, and the cost in time and effort of determining it, 
but until the end of the survey this is unlikely to be the limiting factor in overall precision.

@@

Some of the methods in the challenge accepted target bin edges as an input to their algorithms,
and then used classifiers to assign objects to those bins.  Entrants could select these edges; many followed the example random forest entry and split into equal counts per bin. Other methods optimized the values
of bin edges themselves, by maximizing a metric.  For the runs shown here, these methods mostly
targeted one of the 3x2pt metrics.

@@

One important point to note is that while trained-edge
methods dominate the highest scores for the 3x2pt DETF metric, on which they trained, they
fare much worse on the lensing-only metrics (the methods were not re-trained on the lensing metrics;
we show the metrics for the same assigned bins).  This suggests that the difference between
optimal bins for different science cases is a significant one, and thus that analysis pipelines
should make it straightforward to calibrate multiple configurations and pass them to
subsequent science anlysis; this will probably be true even when comparing models
differing only in their systematic parametrizations.  
This feature is highlighted by the comparison between the two variants
of the \textsc{NeuralNetwork} entry, which differed only in which metric they were targeting;
optimizing one metric does not optimize the other (though the changes between results on the two simulations
could also indicate that this is related to the high-redshift behaviour in the CosmoDC2 simulation
described in \ref{sec:cosmodc2}).

@@

\begin{figure*}
\includegraphics[width=1\linewidth]{results/metric_grid_dc2_griz.pdf}
\caption{Normalized scores for each method, normalized between the score for random assignment (zero) and truth assignment to bins with equal number counts (one).  It can be seen that both equal-number and equal-redshift spacings can be beaten for certain metrics.  In each case, classification is done using griz bands on the CosmoDC2 data set; equivalent plots for Buzzard and riz are shown in \autoref{app:metric_grids}.}
\label{fig:metric_grid_dc2_griz}
\end{figure*}

@@

Several of the methods with optimized target bin edges break this
barrier and achieve very high scores.  This included ZotNet, ZotBin,  and JaxResNet.  The
Stacked Generalization method similarly included manually tuned bin edges to maximize
score for this scenario (and indeed its ensemble combination method could be applied to
combine the other successful methods). The typical 15--20\% improvement in FoM scores when optimizing
nominal bin edges makes clear that re-selecting edges for each science case is worth the time and effort required.

Notably, though, the winning CosmoDC2 \textsc{funbins} method used the simpler approach desribed above. The 
success of this simple and fast solution suggests this as an easy heuristic for well-optimized nominal edges, 
at least for cases where clustering-like metrics are important.  Its relatively lower
scores on the Buzzard metrics, as noted in \autoref{sec:metric-results}, does however further highlight
the importance of science- and sample-specific training.

@@

Finally, we have shown that in general there is a signficant advantage to optimizing
target bins afresh when considering a new science case; this can be a significant boost
to the final constraining power.
Despite this, evenly spaced bins in fiducial co-moving distance have proven
a good general choice, as shown in the winning method in the challenge, {\sc funbins}.

@@

We use a Gaussian estimate for the covariance \citep{takada_jain}; this also incorporates the number of galaxies in the
sample:
\begin{equation}
    \mathrm{Cov}(C^{ij}_\ell, C^{mn}_\ell) = \frac{1}{(2 \ell + 1)\Delta\ell f_\mathrm{sky}}(D^{im}_\ell  D^{jn}_\ell + D^{in}_\ell D^{jm}_\ell)
\end{equation}
where $D^{ij}_\ell = C^{ij}_\ell + N^{ij}_\ell$, $\Delta_\ell$ is the size in $\ell$ of the binned spectra, and we assume an $f_\mathrm{sky}=0.25$.  The noise spectra are:
\begin{align}
N^{\mathrm{clust},ij}_\ell = \delta_{ij} / n_i \\
N^{\mathrm{lens},ij}_\ell = \delta_{ij} \sigma_e^2 / n_i
\end{align}
where $n_i$ is the angular number density of galaxies in bin $i$ (we asssume equally weighted galaxies, and a 
total number density over all bins of 20 galaxies per square arcminute) and $\sigma_e=0.26$, roughly matching that found in previous surveys \citep{des,desy3a,desy3b,kids,hsc}.

\section{Method Details} \label{app:methods}
This appendix details the different methods submitted to the challenge.

@@

In this implementation we chose nominal bin edges by splittng ensuring equal numbers of training set 
objects were in each bin, and then trained the forest to map magnitudes, magnitude errors, and
galaxy sizes to bin indices.  We use the {\sc scikit-learn} implementation of the random forest
\citep{scikit-learn}.

@@

\subsection{ {\sc Autokeras LSTM}}
Autokeras \citep{autokeras} is an AutoML system based on Keras. Given a general neural network
architecture and a dataset, it searches for the best possible detailed configuration for the problem
at hand.

@@

A technical problem, involving conflicting versions of libraries and incompatible hardware,
meant that this method could not be run in the final challenge.

@@

\subsection{ {\sc SimpleSOM} }
The SimpleSOM algorithm submitted by author AHW utilises a self-organising map
\citep[SOM,][]{Kohonen:1982} to perform  a discretisation of the
high-dimensional colour-magnitude-space of the challenge training and reference
datasets.

We utilise a branched version of the {\tt
kohonen}\footnote{\url{https://github.com/ AngusWright/kohonen.git}} package
\citep{Wright/etal:2020b, Wehrens/Kruisselbrink:2018, Wehrens/Lutgarde:2007}
within R \citep{R}, which we integrate with  the challenge's {\tt python}
classes using the {\tt rpy2} interface\footnote{\url{https://rpy2.github.io}}. 
We train a $101 \times 101$ hexagonal-cell SOM with toroidal topology on the
maximal combination of (non-repeated) colours, $z$-band magnitude, and so-called
`band-triplets'. For example, for the `griz' setup, our SOM is trained on the
combination of  $z$, $g-r$, $g-i$, $g-z$, $r-i$, $r-z$, $i-z$, $g-r-(r-i)$,
$r-i-(i-z)$. This combination proved optimal during testing on the DC2 dataset,
as determined by the 3x2pt SNR metric. The training and reference data then
propagated into this trained SOM, producing like-for-like groupings between the
two datasets. We then compute the mean redshift of the training sample sources
within each cell and the number of reference sources per-cell. By rank-ordering
the cells in mean training-sample redshift, we construct the cumulative
distribution function of reference sample source counts, and split cells into
${n_{\rm tomo}}$ equal-N tomographic bins using this function.

In addition to this base functionality, the SimpleSOM method includes a
mechanism for distilling $N_{\rm cell}$ SOM cells into $N_{\rm group}<N_{\rm
cell}$ groups of cells, while maintaining optimal redshift resolution. This is
done by invoking a hierarchical clustering of SOM cells \citep[see Appendix B of
][]{Wright/etal:2020a}, where in this case we combine cells based on the
similarity of the cells' $n(z)$ moments (mean, median, and normalized median absolute
deviation from median) using complete-linkage clustering. The result of this
procedure is the construction of fewer discrete groups of training and reference
data, while not introducing pathological redshift-distribution broadening (as
happens when, e.g., training using a lower resolution SOM and/or clustering
cells based on distances in colour-magnitude space).

@@

The `por que no los dos' algorithm, submitted by author AHW, is a development of the SimpleSOM algorithm, adding the 
additional complexity of also computing template-based photometric redshift point-estimates using BPZ 
\citep{Benitez:2000}. Photo-$z$ point-estimates are derived using the re-calibrated 
template set of \cite{Capak:2004} in combination with 
the Bayesian redshift prior from \cite{Raichoor/etal:2014}, using only the (g)riz bands supplied in the challenge.

@@

\subsection{ {\sc ComplexSOM} } \label{sec:csom}
This method was submitted by authors DA, AHW, AN, EG, BA, and ABr.

The {\tt ComplexSOM} method implements an additional optimization layer on the
methodology used by {\tt SimpleSOM}.

@@

The rationale behind the {\tt ComplexSOM} method is based on the observation
that, while ${\sf C}^{\rm group}_\ell$ is a slow quantity to calculate for a
large number of groups, ${\sf B}$ and therefore ${\sf C}^{\rm bin}_\ell$ are
very fast. Thus, given an initial set of groups characterizing the
distribution of the full sample in colour space, ${\sf C}^{\rm group}_\ell$
can be precomputed once, and used to find the optimal assignment matrix ${\sf
B}$ that maximizes the figure of merit in Equation \ref{eq:fom},
which can be written in terms of ${\sf C}^{\rm bin}_\ell$ alone as

@@

This method was submitted by the author AP.

>>>>

Various recent work have explored general tomographic binning strategies. The general question of optimization
was recently discussed in detail
in \citet{rainbow} using a self-organizing map (SOM) approach to target up to five tomographic bins.  For this
configuration they find that equally spaced redshift bins are a good proxy
for an optimized method, but note, importantly, that we should not be constrained to trying to directly
match our bin selections to specific tomographic bin edges. 
They also highlight the utility of rejecting outlier or hard-to-classify objects from samples, and the ability of SOMs
to do so cleanly.
\citet{taylor18} considered using
fine tomographic bins as an alternative to fully 3D lensing, and found that the strategy of equal numbers
per bin is less effective at high redshift, a result we will echo in this paper. The fine-binning strategy
also allows a set of nulling methods to be applied, offering various advantages in avoiding poorly-understood
regimes \citep{taylor18b,bnt,xcut}.
Most recently, \citet{euclid-nz} studied this question with respect to the Euclid mission, using the \textit{Flagship}
simulation and applying realistic photometric redshift estimation methods to the
sample (we defer this question in this work). They highlight the power of using large numbers of bins,
differences in ideal binnings when
changing the choice of data (the inclusion galaxy galaxy-lensing), and
note the value in discarding galaxies with high redshift uncertainty. We will echo many of these issues below.

@@

\begin{table*}[]
	\begin{tabular}{|l|p{8cm}|c|}
	\hline
	\textbf{Name} & \textbf{Description} & \textbf{Submitters} \\
	\hline
	\textsc{Random} & Random assignment (control method)  & Organizers \\
	\textsc{IBandOnly} & i-band only (control method)  & Organizers \\
	\textsc{RandomForest} & A collection of individual decision trees & Organizers \\
	\textsc{LSTM} & Long Short-Term Memory neural network & CRB, BF, GT, ESC, EJG \\
	\textsc{AutokerasLSTM} & Automaticaly configured LSTM variant. & CRB, BF, GT, ESC, EJG \\
	\textsc{LGBM} & Light Gradient Boosted Machine & CRB, BF, GT, ESC, EJG \\
	\textsc{TCN} & Temporal Convolution Network & CRB, BF, GT, ESC, EJG \\
	\textsc{CNN} & Convolutional Neural Network & CRB, BF, GT, ESC, EJG \\
	\textsc{Ensemble} & Combination of three other network methods & CRB, BF, GT, ESC, EJG \\
	\textsc{SimpleSOM} & Self-Organizing Map & AHW \\
	\textsc{PQNLD} & Extension of {\sc SimpleSom} combining with template-based redshift estimation & AHW \\
	\textsc{UTOPIA} & Nearest-neighbours, optimised in the limit of large representative training sets & AHW \\
	\textsc{ComplexSOM} & Extension of {\sc SimpleSom} with an additional assignment optimization & DA, AHW, AN, EG, BA, ABr \\
	\textsc{GPzBinning} & Gaussian Process redshift estimator and binner & PH \\
	\textsc{JaxCNN/JaxResnet} & Two related CNN-based bin edge optimizers & AP \\
	\textsc{NeuralNetwork1/2} & Dense network optimizing bin assignment for two metrics & FL \\
	\textsc{PCACluster} & Principal Component Analysis of fluxes then optimized clustering & DG \\
	\textsc{ZotBin/ZotNet} & Two related neural network methods with a preprocesing normalizing flow & DK \\
	\textsc{FFNN} & Feed-forward neural network & EN \\
	\textsc{FunBins} & Random forest with various nominal bin edge selection & ABa \\
	\textsc{MLPQNA} & Multi-layer perceptron & SC, MB \\
	\textsc{Stacked Generalization} & Combination of Gradient Boosted Trees classifiers with standard Random Forests (50/50) & JEC\\
	\hline
	\end{tabular}
	\caption{Methods entered into the challenge. The algorithms are described more fully in \autoref{app:methods}.}
	\label{tab:entrants}
\end{table*}

@@

\begin{table*}[]
	\begin{tabular}{|l|llll|llll|}
		\hline
		& \multicolumn{4}{c|}{\textbf{riz}}      & \multicolumn{4}{c|}{\textbf{griz}}                                \\ \hline
		\textbf{Method} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} \\ \hline
		\input{results/tables/table_FOM_DETF_3x2_dc2.tex}
		\hline
	\end{tabular}
	\caption{Values of the 3x2pt DETF ($w_0,w_a$) figure-of-merit achieved by entrant methods on the 
		CosmoDC2 part of the challenge. The horizontal line separates two general approaches to the problem, as discussed in \autoref{sec:train-vs-fix}: those below the line used fixed
		target edges for tomographic bins and optimized the assignment of objects to those bins,
		whereas those above also optimized the target edges themselves.
        Asterisks indicate runs which failed during classification, and dashes indicate runs with
        number counts too small for spectra to be generated.
        }
	\label{tab:cosmodc2}
\end{table*}

@@

\begin{figure}[htbp]
	% \centering
	\includegraphics[width=1\linewidth]{results/nzbinned_dc2_funbins.pdf}
	\caption{The tomographic $n(z)$ values generated by the winning method in the challenge, {\sc FunBins},
		generating nine bins with the CosmoDC2 data.
		The upper panel shows results for the method applied to $riz$ bands, and the middle on $griz$.
		The lower panel is for comparison and is not part of the challenge;
		it illustrates {\sc FunBins} results using all the $ugrizy$ bands.
		The increased spread and overlap of the bins as bands are removed decreases the overall constraining
		power and signal-to-noise.
	}
	\label{fig:funbin_nz}
\end{figure}

No one method dominates the metrics across the different scenarios considered.
Before the challenge began we somewhat arbitrarily chose 
the 9-bin 3x2pt $griz$ CosmoDC2 $w_0-w_a$ metric as the fiducial scenario on which
an overall winning method would be selected.
By this criterion the best method was, by a small margin, \textsc{FunBins},
which used a random forest to assign galaxies to after choosing nominal bin edges that split 
the range spanned
by the sample into equal co-moving distance bins. 
As described in \autoref{sec:funbins},  this splitting was designed
to minimize the shot noise for the angular power in each bin, and this has proven successful. 
Figure \ref{fig:funbin_nz} shows the number density obtained by this method, for both the riz and griz 
scenarios, and an additional post-challenge run using the full Rubin $ugrizy$ bands.
The metrics \textsc{FunBins}
obtained on $ugrizy$ are barely higher than those with $griz$ -- the largest
increase is 8\% for the $\Omega_c-\sigma_8$ metric, with the remainder all 5\% or less.

@@

\begin{figure*}[htbp]
% \centering
\includegraphics[width=0.9\linewidth]{results/nzoverlap_dc2_funbins.pdf}
\caption{The pairwise fractions $f_\mathrm{mis}$ of total objects in redshift regions where multiple tomographic bins contain galaxies, for 
the \textsc{FunBins} tomography method and three choices of bands.  
The $x$- and $y$-axes 
show tomographic bin indices, 
and the colour represents the fraction of objects (compared to the total number in the challenge) that are 
at redshifts where both tomographic bins have members, in bins of width $\delta z=10^{-3}$.  
This metric corresponds to
the fraction of the area that is shaded in two colours in Figure \ref{fig:funbin_nz}.
The total-overlap figure gives the total fraction of objects
that are ambiguous over all bins, and is the sum of all the pairwise values.}
\label{fig:overlap}
\end{figure*}

\textsc{FunBins} seems to be a good and stable general method, though several other methods perform as well
or slightly better in other scenarios.  Notably, many methods reach the same plateau in lensing-only metrics,
suggesting this is somewhat easier to optimize; this is consistent with the general reason that photometric
redshifts suffice for lensing: the broad lensing kernels are less sensitive to small changes.

@@

\subsection{Non-assignment}
The challenge did not require entrants to assign all galaxies to bins; aside from the impact of reduced final sample size on the metrics, no additional penalty was applied for discarding galaxies entirely.  Several methods
took advantage of this.  The \textsc{ComplexSOM} method (see \autoref{sec:csom}) found
it did not improve scores, but as noted in \autoref{sec:pqnld} this may be an artifact
of the high density of the training data, and so could be explored in future challenges\footnote{In real data or more realistic simulations discarding galaxies has more complicated effects than simply decreasing sample size as it does here - removing objects with broad $p(z)$ can narrow the final $n(z)$, not just lower it. This was not a factor in this challenge because we use true redshifts to calculate our metrics.}. See \citet{euclid-nz} for a discussion of this topic in the Euclid context.

@@

The different metrics are generally well-correlated (this also holds true for the metrics not 
shown here), especially at the highest values of the metrics representing the most successful
methods. This is particularly so when comparing the Buzzard and CosmoDC2 metrics, which is encouraging 
as it implies these methods are broadly stable to reasonable changes in the underlying galaxy
distribution, and hence on the survey properties.  The largest outlier comes from the 
{\sc NeuralNetwork} entry.

@@

The value of the extra FoM that could be gained by adding the band should be weighed against 
the challenge of high-accuracy
PSF measurement for this band, and the cost in time and effort of determining it, 
but until the end of the survey this is unlikely to be the limiting factor in overall precision.

Some methods perform worse when adding the $g$-band, which we ascribe to reduced performance
in their machine learning algorithms when increasing the dimensionality of their inputs; further tuning
of their hyperparameters could perhaps alleviate the issue.

@@

Some of the methods in the challenge accepted target bin edges as an input to their algorithms,
and then used classifiers to assign objects to those bins.  Entrants could select these edges; many followed the example random forest entry and split into equal counts per bin; some like \textsc{FunBins} used alternative
criteria. Other methods optimized the values
of bin edges themselves, by maximizing a metric.  For the runs shown here, these methods mostly
targeted one of the 3x2pt metrics.

@@

One important point to note is that while trained-edge
methods dominate the highest scores for the 3x2pt DETF metric, on which they trained, they
fare much worse on the lensing-only metrics (the methods were not re-trained on the lensing metrics;
we show the metrics for the same assigned bins).  This suggests that the difference between
optimal bins for different science cases is a significant one, and thus that analysis pipelines
should make it straightforward to calibrate multiple configurations and pass them to
subsequent science analysis; this will probably be true even when comparing models
differing only in their systematic parametrizations.  
This feature is highlighted by the comparison between the two variants
of the \textsc{NeuralNetwork} entry, which differed only in which metric they were targeting;
optimizing one metric does not optimize the other (though the changes between results on the two simulations
could also indicate that this is related to the high-redshift behaviour in the CosmoDC2 simulation
described in \ref{sec:cosmodc2}).

@@

\begin{figure*}
\includegraphics[width=1\linewidth]{results/metric_grid_dc2_griz.pdf}
\caption{Normalized scores for each method, normalized between the score for random assignment (zero) and truth assignment to bins with equal number counts (one).  It can be seen that both equal-number and equal-redshift spacings can be beaten for certain metrics.  In each case, classification is done using griz bands on the CosmoDC2 data set; equivalent plots for Buzzard and riz are shown in \autoref{app:metric_grids}.
\emph{ww} indicates lensing-only metrics, \emph{gg} indicates clustering only, and \emph{3x2pt} their combination.
The three metrics are defined in \autoref{sec:metrics}.
}
\label{fig:metric_grid_dc2_griz}
\end{figure*}

@@

Several of the methods with optimized target bin edges break this
barrier and achieve very high scores.  This included \textsc{ZotNet}, \textsc{ZotBin}, and \textsc{JaxResNet}.  The
\textsc{Stacked Generalization} method similarly included manually tuned bin edges to maximize
score for this scenario (and indeed its ensemble combination method could be applied to
combine the other successful methods). The typical 15--20\% improvement in FoM scores when optimizing
nominal bin edges makes clear that re-selecting edges for each science case is worth the time and effort required.

Notably, though, the winning CosmoDC2 \textsc{FunBins} method used the simpler approach desribed above. The 
success of this simple and fast solution suggests this as an easy heuristic for well-optimized nominal edges, 
at least for cases where clustering-like metrics are important.  Its relatively lower
scores on the Buzzard metrics, as noted in \autoref{sec:metric-results}, does however further highlight
the importance of science- and sample-specific training.

@@

Finally, we have shown that in general there is a signficant advantage to optimizing
target bins afresh when considering a new science case; this can be a significant boost
to the final constraining power.
Despite this, evenly spaced bins in fiducial co-moving distance have proven
a good general choice, as shown in the winning method in the challenge, {\sc FunBins}.

@@

We use a Gaussian estimate for the covariance \citep{takada_jain}; this also incorporates the number of galaxies in the
sample:
\begin{equation}
    \mathrm{Cov}(C^{ij}_\ell, C^{mn}_\ell) = \frac{1}{(2 \ell + 1)\Delta\ell f_\mathrm{sky}}(D^{im}_\ell  D^{jn}_\ell + D^{in}_\ell D^{jm}_\ell)
\end{equation}
where $D^{ij}_\ell = C^{ij}_\ell + N^{ij}_\ell$, $\Delta\ell$ is the size in $\ell$ of the binned spectra, and we assume an $f_\mathrm{sky}=0.25$.  The noise spectra are:
\begin{align}
N^{\mathrm{clust},ij}_\ell = \delta_{ij} / n_i \\
N^{\mathrm{lens},ij}_\ell = \delta_{ij} \sigma_{e,\mathrm{tot}}^2 / n_i
\end{align}
where $n_i$ is the angular number density of galaxies in bin $i$ (we asssume equally weighted galaxies, and a 
total number density over all bins of 20 galaxies per square arcminute) and $\sigma_{ei}=0.26$ per shear component, giving $\sigma_{e,\mathrm{tot}} = 0.37$, roughly matching that found in previous surveys \citep{des,desy3a,desy3b,kids,hsc}.

This model does not include a number of theoretical effects that will be important in real data, including
intrinsic alignments, non-Gaussian covariance, and a range of effects on spectra at small scales, like non-linear 
bias and beyond-Limber effects. Collectively these effects mean that our tests are exaggerating the constraining
power of small-scale signals, and increasing all our metrics (especially as our $\ell$ range is fixed with $z$).  
Although this unrealistic aspect mostly affects algorithms equally, the relation between redshift, angle, and physical scale means that the exaggeration is especially strong for low redshift, and so it
could plausibly favour methods that are better in that range.

\section{Method Details} \label{app:methods}
This appendix details the different methods submitted to the challenge.
Methods described as using optimized bin edges are those that appear as orange solid
lines in \autoref{fig:edges}; those without this note appear as dashed blue lines there.

@@

In this implementation we chose nominal bin edges by splitting ensuring equal numbers of training set 
objects were in each bin, and then trained the forest to map magnitudes, magnitude errors, and
galaxy sizes to bin indices.  We use the {\sc scikit-learn} implementation of the random forest
\citep{scikit-learn}.

@@

\subsection{ {\sc AutokerasLSTM}}
Autokeras \citep{autokeras} is an AutoML system based on Keras. Given a general neural network
architecture and a dataset, it searches for the best possible detailed configuration for the problem
at hand.

@@

@@

\subsection{ {\sc SimpleSOM} }
The \textsc{SimpleSOM} algorithm submitted by author AHW utilises a self-organising map
\citep[SOM,][]{Kohonen:1982} to perform  a discretisation of the
high-dimensional colour-magnitude-space of the challenge training and reference
datasets.

We utilise a branched version of the {\sc
kohonen}\footnote{\url{https://github.com/ AngusWright/kohonen.git}} package
\citep{Wright/etal:2020b, Wehrens/Kruisselbrink:2018, Wehrens/Lutgarde:2007}
within R \citep{R}, which we integrate with  the challenge's python
classes using the {\sc rpy2} interface\footnote{\url{https://rpy2.github.io}}. 
We train a $101 \times 101$ hexagonal-cell SOM with toroidal topology on the
maximal combination of (non-repeated) colours, $z$-band magnitude, and so-called
`band-triplets'. For example, for the `griz' setup, our SOM is trained on the
combination of  $z$, $g-r$, $g-i$, $g-z$, $r-i$, $r-z$, $i-z$, $g-r-(r-i)$,
$r-i-(i-z)$. This combination proved optimal during testing on the DC2 dataset,
as determined by the 3x2pt SNR metric. The training and reference data then
propagated into this trained SOM, producing like-for-like groupings between the
two datasets. We then compute the mean redshift of the training sample sources
within each cell and the number of reference sources per-cell. By rank-ordering
the cells in mean training-sample redshift, we construct the cumulative
distribution function of reference sample source counts, and split cells into
${n_{\rm tomo}}$ equal-N tomographic bins using this function.

In addition to this base functionality, the \textsc{SimpleSOM} method includes a
mechanism for distilling $N_{\rm cell}$ SOM cells into $N_{\rm group}<N_{\rm
cell}$ groups of cells, while maintaining optimal redshift resolution. This is
done by invoking a hierarchical clustering of SOM cells \citep[see Appendix B of
][]{Wright/etal:2020a}, where in this case we combine cells based on the
similarity of the cells' $n(z)$ moments (mean, median, and normalized median absolute
deviation from median) using complete-linkage clustering. The result of this
procedure is the construction of fewer discrete groups of training and reference
data, while not introducing pathological redshift-distribution broadening (as
happens when, e.g., training using a lower resolution SOM and/or clustering
cells based on distances in colour-magnitude space).

@@

The `por que no los dos' algorithm, submitted by author AHW, is a development of the \textsc{SimpleSOM} algorithm, adding the 
additional complexity of also computing template-based photometric redshift point-estimates using BPZ 
\citep{Benitez:2000}. Photo-$z$ point-estimates are derived using the re-calibrated 
template set of \cite{Capak:2004} in combination with 
the Bayesian redshift prior from \cite{Raichoor/etal:2014}, using only the (g)riz bands supplied in the challenge.

@@

\subsection{ {\sc ComplexSOM} } \label{sec:csom}
This method was submitted by authors DA, AHW, AN, EG, BA, and ABr. 
It used optimized bin edges.

The {\sc ComplexSOM} method implements an additional optimization layer on the
methodology used by {\sc SimpleSOM}.

@@

The rationale behind the {\sc ComplexSOM} method is based on the observation
that, while ${\sf C}^{\rm group}_\ell$ is a slow quantity to calculate for a
large number of groups, ${\sf B}$ and therefore ${\sf C}^{\rm bin}_\ell$ are
very fast. Thus, given an initial set of groups characterizing the
distribution of the full sample in colour space, ${\sf C}^{\rm group}_\ell$
can be precomputed once, and used to find the optimal assignment matrix ${\sf
B}$ that maximizes the figure of merit in Equation \ref{eq:fom},
which can be written in terms of ${\sf C}^{\rm bin}_\ell$ alone as

@@

These method were submitted by the author AP.
They used optimized bin edges.
